---
title: "december"
author: "Gabrielle"
date: '2022-12-13'
output: 
  html_document:
    code_folding: hide
---
# Socio-demographic Patterning of Microplastic Burden in the U.S.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning = FALSE,results=FALSE)
library(tidyverse)
library(here)
library(ggpubr) # for some graphic applications that extend ggplot2
library(janitor)
library(broom) # used to make tables
library(knitr) # used to make table
library(car) # has leveneTest 
library(foreign) # to read in NHANES data
library(rstanarm) # for the model fitting
library(jtools) # Load jtools, for forest plots
library(sandwich) # needed for robust standard errors in forest plot
library(huxtable) # needed to be able to export table of forest plot values
library(broom.mixed) # used in making tables
library(visdat)
library(ggplot2)
```

## Load in the data: years 2001-2016
- demographics file
- phthalates, urine file

downloaded from CDC National Health and Examination Survey, [NHANES Questionnaires, Datasets, and Related Documentation](https://wwwn.cdc.gov/nchs/nhanes/Default.aspx)

```{r}
# 2001-2002
demographics01 <- read.xport(here("data","DEMO_B_2001_2002.XPT"))
phthalates01 <- read.xport(here("data","PHPYPA_B_2001_2002.XPT"))
nhanes01 <- merge(phthalates01,demographics01,all.x=T)

# 2003-2004
demographics03 <- read.xport(here("data","DEMO_C_2003_2004.XPT"))
phthalates03 <- read.xport(here("data","L24PH_C_2003_2004.XPT"))
nhanes03 <- merge(phthalates03,demographics03,all.x=T)

# 2005-2006
demographics05 <- read.xport(here("data","DEMO_D_2005_2006.XPT"))
phthalates05 <- read.xport(here("data","PHTHTE_D_2005_2006.XPT"))
nhanes05 <- merge(phthalates05,demographics05,all.x=T)

# 2007-2008
demographics07 <- read.xport(here("data","DEMO_E_2007_2008.XPT"))
phthalates07 <- read.xport(here("data","PHTHTE_E_2007_2008.XPT"))
nhanes07 <- merge(phthalates07,demographics07,all.x=T)

# 2009-2010
demographics09 <- read.xport(here("data","DEMO_F_2009_2010.XPT"))
phthalates09 <- read.xport(here("data","PHTHTE_F_2009_2010.XPT"))
nhanes09 <- merge(phthalates09,demographics09,all.x=T)

# 2011-2012
demographics11 <- read.xport(here("data","DEMO_G_2011_2012.XPT"))
phthalates11 <- read.xport(here("data","PHTHTE_G_2011_2012.XPT"))
nhanes11 <- merge(phthalates11,demographics11,all.x=T)

# 2013-2014
demographics13 <- read.xport(here("data","DEMO_H_2013_2014.XPT"))
phthalates13 <- read.xport(here("data","PHTHTE_H_2013_2014.XPT"))
nhanes13 <- merge(phthalates13,demographics13,all.x=T)

# 2015-2016
demographics15 <- read.xport(here("data","DEMO_I_2015_2016.XPT"))
phthalates15 <- read.xport(here("data","PHTHTE_I_2015_2016.XPT"))
nhanes15 <- merge(phthalates15,demographics15,all.x=T)

## add a column with the year to each data file:

nhanes01$year <- 2001
nhanes03$year <- 2003
nhanes05$year <- 2005
nhanes07$year <- 2007
nhanes09$year <- 2009
nhanes11$year <- 2011
nhanes13$year <- 2013
nhanes15$year <- 2015

## choose the variables to select from each data file:
var<-c('year','DMDEDUC3','DMDEDUC2','DMDHREDU','DMDHSEDU','RIDAGEYR','RIAGENDR','RIDRETH1','INDFMPIR','DMDYRSUS','DMDCITZN','URXMEP')
nhanes01s<-subset(nhanes01,select=var)
nhanes03s<-subset(nhanes03,select=var)
nhanes05s<-subset(nhanes05,select=var)
nhanes07s<-subset(nhanes07,select=var)
nhanes09s<-subset(nhanes09,select=var)
nhanes11s<-subset(nhanes11,select=var)
nhanes13s<-subset(nhanes13,select=var)
nhanes15s<-subset(nhanes15,select=var)

## create one dataset for NHANES demographics and phthalates, year 2001-2016
nhanes <- rbind(nhanes01s,nhanes03s,nhanes05s,nhanes07s,nhanes09s,nhanes11s,nhanes13s,nhanes15s)

## create a data file so that I don't need to load, clean, and select the data again
# has all demographics and URXMEP phthalate for NHANES year 2001-2016
write.csv(nhanes, "nhanes_december22.csv")

```


## Visually inspect the data:
After inspecting the phthalate data in its raw form and the log form, we determined it is necessary to log transform the data in order to visualize it.

```{r}
hist(nhanes$URXMEP)
hist(log(nhanes$URXMEP))

nhanes$log_monoethyl = log(nhanes$URXMEP)
```

## Re-categorize the data:
Original:
- year = year the observation took place *this was added by me when I merged the years of data
- DMDEDUC3 = the highest grade level of education completed by participants 6-19 y.o.
- DMDEDUC2 = the highest grade/level of education completed by participants 20 years and older
- DMDHREDU = the household reference person's education level
- DMDHSEDU = the household reference person's spouse's education level
- RIDAGEYR = ages in years at screening
- RIAGENDR = gender
- RIDRETH1 = race/Hispanic origin
- INDFMPIR = ratio of family income to poverty
- DMDYRSUS = length of time in U.S.
- DMDCITZN = citizenship status
- URXMEP = mono-ethyl phthalate (ng/mL)

## Household reference person's education level
```{r}
hist(nhanes$DMDHREDU)

# add "edu_ref" to "nhanes" 
nhanes$edu_ref = factor(ifelse(nhanes$DMDHREDU %in% 1:4,"partial college and below",
                         ifelse(nhanes$DMDHREDU==5,"college and beyond",NA)),
                  levels=c("partial college and below","college and beyond"))
summary(nhanes$edu_ref)

```

## Household reference person's spouse's education level
```{r}
hist(nhanes$DMDHSEDU)

# add "edu_spouse_ref" to "nhanes" 
nhanes$edu_spouse_ref = factor(ifelse(nhanes$DMDHSEDU %in% 1:4,"partial college and below",
                         ifelse(nhanes$DMDHSEDU==5,"college and beyond",NA)),
                  levels=c("partial college and below","college and beyond"))
summary(nhanes$edu_spouse_ref)
```

## Highest grade level of education completed by participants 6-19 y.o.
```{r}
hist(nhanes$DMDEDUC3)

# add "edu_child_participants" to "nhanes" 
nhanes$edu_child_participants = factor(ifelse(nhanes$DMDEDUC3 %in% 0:8,"primary",
                         ifelse(nhanes$DMDEDUC3 %in% 9:15,"secondary",NA)),
                  levels=c("primary","secondary"))
summary(nhanes$edu_child_participants)

## not a good variable d/t 14,755 "refused," "don't know," "missing,
```


## Highest grade/level of education completed by participants 20 years and older
```{r}
hist(nhanes$DMDEDUC2)

# add "edu_adult_participants" to "nhanes" 
nhanes$edu_adult_participants = factor(ifelse(nhanes$DMDEDUC2 ==1, "less than 9th grade",
                                       ifelse(nhanes$DMDEDUC2 ==2, "9-11th grade",
                                       ifelse(nhanes$DMDEDUC2 ==3, "high school grad/GED", 
                                       ifelse(nhanes$DMDEDUC2 ==4, "some college or AA",
                                       ifelse(nhanes$DMDEDUC2 ==5, "college grad or above", NA))))))
summary(nhanes$edu_adult_participants)

```


## Age in years at screening
Based on the summary statistics of the variable "RIDAGEYR": the youngest participant is 3 years old, the oldest participant is 85 years old, and the median age is 31 years old. 

```{r}
hist(nhanes$RIDAGEYR)
summary(nhanes$RIDAGEYR)


# add "age" to "nhanes" 
nhanes$age = factor(ifelse(nhanes$RIDAGEYR >=65,"older adult",
                    ifelse(nhanes$RIDAGEYR >25 & nhanes$RIDAGEYR<=64,"middle-aged",
                    ifelse(nhanes$RIDAGEYR >=19 & nhanes$RIDAGEYR<=25,"young adult",
                    ifelse(nhanes$RIDAGEYR <=18,"child",NA)))),
             levels=c("child","young adult", "middle-aged", "older adult"))
summary(nhanes$age)



```

## Gender
```{r}
hist(nhanes$RIAGENDR)

# add "gender" to "nhanes"
nhanes$gender = factor(ifelse(nhanes$RIAGENDR ==1, "male",
                       ifelse(nhanes$RIAGENDR ==2, "female", NA)),
                levels=c("male", "female"))
summary(nhanes$gender)

```

## Race/ethnicity
```{r}
hist(nhanes$RIDRETH1)

# add "race/ethnicity" to "nhanes"
nhanes$ethnicity = factor(ifelse(nhanes$RIDRETH1 ==1, "Mexican American",
                       ifelse(nhanes$RIDRETH1 ==2, "Other Hispanic",
                       ifelse(nhanes$RIDRETH1 ==3, "Non-Hispanic White", 
                       ifelse(nhanes$RIDRETH1 ==4, "Non-Hispanic Black",
                       ifelse(nhanes$RIDRETH1 ==5, "Other or Multi", NA))))),
                levels=c("Non-Hispanic White", "Non-Hispanic Black", "Mexican American", "Other Hispanic", "Other or Multi"))
summary(nhanes$ethnicity)

```

## Citizenship Status
```{r}
hist(nhanes$DMDCITZN)

# add "citizenship status" to "nhanes"
nhanes$citizenship = factor(ifelse(nhanes$DMDCITZN ==1, "birth or naturalization",
                       ifelse(nhanes$DMDCITZN ==2, "not U,S, citizen", NA)),
                levels=c("birth or naturalization", "not U,S, citizen"))
summary(nhanes$citizenship)
```

## Length of time in the U.S.
```{r}
hist(nhanes$DMDYRSUS)

# add length of time in the U.S. to "nhanes"
nhanes$years_in_US = factor(ifelse(nhanes$DMDYRSUS ==1, "less than 1 year",
                            ifelse(nhanes$DMDYRSUS ==2, "1 year or more, but less than 5 years",
                            ifelse(nhanes$DMDYRSUS ==3, "5 years or more, but less than 10 years",
                            ifelse(nhanes$DMDYRSUS ==4, "10 years or more, but less than 15 years",
                            ifelse(nhanes$DMDYRSUS ==5, "15 years or more, but less than 20 years",
                            ifelse(nhanes$DMDYRSUS ==6, "20 years or more, but less than 30 years",
                            ifelse(nhanes$DMDYRSUS ==7, "30 years or more, but less than 40 years",
                            ifelse(nhanes$DMDYRSUS ==8, "40 years or more, but less than 50 years",
                            ifelse(nhanes$DMDYRSUS ==9, "50 years or more", NA))))))))))
summary(nhanes$years_in_US)

## not a very good variable, d/t 18,035 "refused," "don't know", or "missing" 
```

## Ratio of Family Income to Poverty Level
```{r}
hist(nhanes$INDFMPIR)

# add ratio of family income to poverty level
nhanes$income_to_poverty_ratio = factor(ifelse(nhanes$INDFMPIR < 1,"at poverty threshold",
                                        ifelse(nhanes$INDFMPIR >=1 & nhanes$INDFMPIR <2, "family income 2x poverty threshold",
                                        ifelse(nhanes$INDFMPIR >=2 & nhanes$INDFMPIR <3, "family income 3x poverty threshold",
                                        ifelse(nhanes$INDFMPIR >=3 & nhanes$INDFMPIR <4, "family income 4x poverty threshold",
                                        ifelse(nhanes$INDFMPIR >=4 & nhanes$INDFMPIR <5, "family income 5x poverty threshold",
                                        ifelse(nhanes$INDFMPIR ==5, "family income more than 5x poverty threshold", NA)))))),
                                 levels=c("at poverty threshold", "family income 2x poverty threshold", "family income 3x poverty threshold", "family income 4x poverty threshold", "family income 5x poverty threshold", "family income more than 5x poverty threshold"))

summary(nhanes$income_to_poverty_ratio)
```

## Boxplots
```{r}
# Household reference person education level and Mono-ethyl phthalate
box_edu_ref <- ggplot(data = nhanes, aes(x=log_monoethyl, y=edu_ref, fill=edu_ref)) + 
  scale_fill_brewer(palette="PuBuGn") +
  geom_boxplot() +
  theme(text = element_text(size=12)) +
  xlab("(logged) Mono-Ethyl Phthalate Level (ng/mL)") +
  ylab("Household Reference Person's Education Level") +
ggtitle("Household Ref. Person's Education Level and Logged Phthalate Level")

box_edu_ref

# Family Income and Mono-ethyl phthalate
box_poverty <- ggplot(data = nhanes, aes(x=log_monoethyl, y=income_to_poverty_ratio, fill=income_to_poverty_ratio)) + 
  scale_fill_brewer(palette="PuBuGn") +
  geom_boxplot() +
  theme(text = element_text(size=12)) +
  xlab("(logged) Mono-Ethyl Phthalate Level (ng/mL)") +
  ylab("Ratio of Participant's Family Income to Poverty") +
  ggtitle("Poverty and Logged Phthalate Level")

box_poverty

# Age and Mono-ethyl phthalate
box_age <- ggplot(data = nhanes, aes(x=log_monoethyl, y=age, fill=age)) + 
  scale_fill_brewer(palette="PuBuGn") +
  geom_boxplot() +
  theme(text = element_text(size=12)) +
  xlab("(logged) Mono-Ethyl Phthalate Level (ng/mL)") +
  ylab("Age of Participant") +
  ggtitle("Age and Logged Phthalate Level")

box_age
```


## Suite of Models
```{r}
# multivariate regression analysis of socio-demographic variables and phthalates
modela <- lm(log_monoethyl ~ edu_ref + age + gender + ethnicity + income_to_poverty_ratio +
                  citizenship, data = nhanes)
summary(modela)

# take out gender
modelb <- lm(log_monoethyl ~ edu_ref + age + ethnicity + income_to_poverty_ratio +
                  citizenship, data = nhanes)
summary(modelb)

# take out gender, citizenship
modelc <- lm(log_monoethyl ~ edu_ref + age + ethnicity + income_to_poverty_ratio, data = nhanes)
summary(modelc)


```


## Bayesian Information Criterion (BIC)
The model that performs best is Model B, a simple linear regression showing the logged mono-ethyl phthalate as a function of the reference person's education level, the participant's age, the participant's ethnicity, the participant's family income to poverty ratio, and the participant's citizenship status.

The delta BIC for Model C is 72.136, which is far beyond the BIC <7 threshold, so it is a very unlikely model, and therefore, we will dismiss Model C. This reinforces the idea that citizenship status matters substantially to the question of determining the logged phthalate level in individuals.

Model A and Model B are quite similar - the only difference is that Model B does not include gender. Model A, which is the model performing second best, has a delta BIC = 6.590 (and still under the BIC < 7 threshold). 

From Model B, we can determine --- EQUATION?



```{r, results = TRUE}
BIC_list <- c(BIC(modela), BIC(modelb), BIC(modelc))

model_output <- rbind(data.frame(glance(modela)), data.frame(glance(modelb)), data.frame(glance(modelc))) %>% select(BIC)

model_output <- mutate(model_output, delta.BIC = BIC-min(BIC_list))
model_output$model <- c("Model A", "Model B", "Model C")
model_output <- model_output[,c("model", "BIC", "delta.BIC")]

kable(model_output, format = "markdown", digits = 3, caption = "BIC, and Delta.BIC for the models. Delta BIC > 7 indicates models that should be dismissed from further consideration.")
```


## Forest Models

### Forest model for Model A:
```{r}
data(nhanes) # Telling R we want to use this data

modela <- lm(log_monoethyl ~ edu_ref + age + gender + ethnicity + income_to_poverty_ratio +
                  citizenship, data = nhanes)
summ(modela)

summ(modela, robust = "HC1") #robust standard errors 

summ(modela, confint = TRUE, digits = 3) #In many cases, you’ll learn more by looking at confidence intervals than p-values. You can request them from summ. default is 95% CIs

summ(modela, confint = TRUE, pvals = FALSE) #DROP the p values all together

# THE GRAPH
plot_summs(modela)
plot_summs(modela, robust = TRUE)
plot_summs(modela, inner_ci_level = .9)

# plot coefficient uncertainty as normal distributions
plot_summs(modela, plot.distributions = TRUE, inner_ci_level = .9)

# table output for Word and RMarkdown documents
## error is in the parenthesis
export_summs(modela, scale = TRUE)

# confidence intervals instead of standard errors
export_summs(modela, scale = TRUE,
             error_format = "[{conf.low}, {conf.high}]")

```
#### renaming variables in the forest plot
```{r}
forest1 <- plot_summs(
        point.size = 3,
        fontsize=8,
        colors = "darkseagreen4",
        modela, coefs = c("Household Education College and Beyond
                                 Partial College and Below (ref)" = "edu_refcollege and beyond", 
                          
                                 "Age: Young Adult
                                 Child (ref)" = "ageyoung adult",
                          
                                 "Age: Middle-aged
                                 Age:Child (ref)" = "agemiddle-aged",
                          
                                 "Age: Older Adult
                                 Age: Child (ref)" = "ageolder adult",
                          
                                 "Gender: Female
                                 Gender: Male (ref)" = "genderfemale",
                          
                                 "Ethnicity: Non-Hispanic Black
                                 Non-Hispanic White (ref)" = "ethnicityNon-Hispanic Black",
                          
                                 "Ethnicity: Mexican American
                                 Non-Hispanic White (ref)" = "ethnicityMexican American",
                          
                                 "Ethnicity: Other Hispanic
                                 Non-Hispanic White (ref)" = "ethnicityOther Hispanic",
                          
                                 "Ethnicity: Other or Multi
                                 Non-Hispanic White (ref)" = "ethnicityOther or Multi",
                          
                                 "Family Income to Poverty Ratio: 2x Poverty threshold
                                 At poverty threshold (ref)" = "income_to_poverty_ratiofamily income 2x poverty threshold",
                          
                                 "Family Income to Poverty Ratio: 3x Poverty threshold
                                 At poverty threshold (ref)" = "income_to_poverty_ratiofamily income 3x poverty threshold",
                          
                                 "Family Income to Poverty Ratio: 4x Poverty threshold
                                 At poverty threshold (ref)" = "income_to_poverty_ratiofamily income 4x poverty threshold",
                          
                                 "Family Income to Poverty Ratio: 5x Poverty threshold
                                 At poverty threshold (ref)" = "income_to_poverty_ratiofamily income 5x poverty threshold",
                          
                                 "Family Income to Poverty Ratio: more than 5x Poverty threshold
                                 At poverty threshold (ref)" = "income_to_poverty_ratiofamily income more than 5x poverty threshold",
                          
                                 "Citizenship Status: Not U.S. Citizen
                                 U.S. Citizen by birth or naturalization (ref)" = "citizenshipnot U,S, citizen"),
        
                          scale = TRUE, robust = TRUE)

forest1
```


### Forest model for Model B:
```{r}
data(nhanes) # Telling R we want to use this data

modelb <- lm(log_monoethyl ~ edu_ref + age + ethnicity + income_to_poverty_ratio +
                  citizenship, data = nhanes)
summ(modelb)

summ(modelb, robust = "HC1") #robust standard errors 

summ(modelb, confint = TRUE, digits = 3) #In many cases, you’ll learn more by looking at confidence intervals than p-values. You can request them from summ. default is 95% CIs

summ(modelb, confint = TRUE, pvals = FALSE) #DROP the p values all together

# THE GRAPH
plot_summs(modelb)
plot_summs(modelb, robust = TRUE)
plot_summs(modelb, inner_ci_level = .9)

# plot coefficient uncertainty as normal distributions
plot_summs(modelb, plot.distributions = TRUE, inner_ci_level = .9)

# table output for Word and RMarkdown documents
## error is in the parenthesis
export_summs(modelb, scale = TRUE)

# confidence intervals instead of standard errors
export_summs(modelb, scale = TRUE,
             error_format = "[{conf.low}, {conf.high}]")

```
#### renaming variables in the forest plot 2
```{r}
forest2 <- plot_summs(
        point.size = 3,
        fontsize=8,
        colors = "darkseagreen3",
        modelb, coefs = c("Household Education College and Beyond
                                 Partial College and Below (ref)" = "edu_refcollege and beyond", 
                          
                                 "Age: Young Adult
                                 Child (ref)" = "ageyoung adult",
                          
                                 "Age: Middle-aged
                                 Age:Child (ref)" = "agemiddle-aged",
                          
                                 "Age: Older Adult
                                 Age: Child (ref)" = "ageolder adult",
                          
                                 "Ethnicity: Non-Hispanic Black
                                 Non-Hispanic White (ref)" = "ethnicityNon-Hispanic Black",
                          
                                 "Ethnicity: Mexican American
                                 Non-Hispanic White (ref)" = "ethnicityMexican American",
                          
                                 "Ethnicity: Other Hispanic
                                 Non-Hispanic White (ref)" = "ethnicityOther Hispanic",
                          
                                 "Ethnicity: Other or Multi
                                 Non-Hispanic White (ref)" = "ethnicityOther or Multi",
                          
                                 "Family Income to Poverty Ratio: 2x Poverty threshold
                                 At poverty threshold (ref)" = "income_to_poverty_ratiofamily income 2x poverty threshold",
                          
                                 "Family Income to Poverty Ratio: 3x Poverty threshold
                                 At poverty threshold (ref)" = "income_to_poverty_ratiofamily income 3x poverty threshold",
                          
                                 "Family Income to Poverty Ratio: 4x Poverty threshold
                                 At poverty threshold (ref)" = "income_to_poverty_ratiofamily income 4x poverty threshold",
                          
                                 "Family Income to Poverty Ratio: 5x Poverty threshold
                                 At poverty threshold (ref)" = "income_to_poverty_ratiofamily income 5x poverty threshold",
                          
                                 "Family Income to Poverty Ratio: more than 5x Poverty threshold
                                 At poverty threshold (ref)" = "income_to_poverty_ratiofamily income more than 5x poverty threshold",
                          
                                 "Citizenship Status: Not U.S. Citizen
                                 U.S. Citizen by birth or naturalization (ref)" = "citizenshipnot U,S, citizen"),
        
                          scale = TRUE, robust = TRUE)

forest2
```


## Visualize Missing Data
```{r}
vis_miss(nhanes, sort_miss = TRUE)
```







## References:
- add a reference to demonstrate rationale for categorizing each variable the way that I do (ex - college completion and above and partial college and below)
- use resources from GEOG 227


