---
title: "Adding Weights to NHANES"
author: "Gabrielle"
date: '2023-02-07'
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning = FALSE,results=FALSE)
library(tidyverse)
library(here)
library(ggpubr) # for some graphic applications that extend ggplot2
library(janitor)
library(broom) # used to make tables
library(knitr) # used to make table
library(car) # has leveneTest 
library(foreign) # to read in NHANES data
library(rstanarm) # for the model fitting
library(jtools) # Load jtools, for forest plots
library(sandwich) # needed for robust standard errors in forest plot
library(huxtable) # needed to be able to export table of forest plot values
library(broom.mixed) # used in making tables
library(visdat)
library(ggplot2)

library(haven) # for reading SAS XPT file from NHANES website
library(survey) # for using survey weights
library(dplyr) # for data wrangling

```

## LOAD THE DATASET 

Years 2001-2016
- demographics file
- phthalates, urine file

Downloaded from CDC National Health and Examination Survey, [NHANES Questionnaires, Datasets, and Related Documentation](https://wwwn.cdc.gov/nchs/nhanes/Default.aspx)


```{r}
# 2001-2002
demographics01 <- read.xport(here("data","DEMO_B_2001_2002.XPT"))
phthalates01 <- read.xport(here("data","PHPYPA_B_2001_2002.XPT"))
nhanes01 <- merge(phthalates01,demographics01,all.x=T)

# 2003-2004
demographics03 <- read.xport(here("data","DEMO_C_2003_2004.XPT"))
phthalates03 <- read.xport(here("data","L24PH_C_2003_2004.XPT"))
nhanes03 <- merge(phthalates03,demographics03,all.x=T)

# 2005-2006
demographics05 <- read.xport(here("data","DEMO_D_2005_2006.XPT"))
phthalates05 <- read.xport(here("data","PHTHTE_D_2005_2006.XPT"))
nhanes05 <- merge(phthalates05,demographics05,all.x=T)

# 2007-2008
demographics07 <- read.xport(here("data","DEMO_E_2007_2008.XPT"))
phthalates07 <- read.xport(here("data","PHTHTE_E_2007_2008.XPT"))
nhanes07 <- merge(phthalates07,demographics07,all.x=T)

# 2009-2010
demographics09 <- read.xport(here("data","DEMO_F_2009_2010.XPT"))
phthalates09 <- read.xport(here("data","PHTHTE_F_2009_2010.XPT"))
nhanes09 <- merge(phthalates09,demographics09,all.x=T)

# 2011-2012
demographics11 <- read.xport(here("data","DEMO_G_2011_2012.XPT"))
phthalates11 <- read.xport(here("data","PHTHTE_G_2011_2012.XPT"))
nhanes11 <- merge(phthalates11,demographics11,all.x=T)

# 2013-2014
demographics13 <- read.xport(here("data","DEMO_H_2013_2014.XPT"))
phthalates13 <- read.xport(here("data","PHTHTE_H_2013_2014.XPT"))
nhanes13 <- merge(phthalates13,demographics13,all.x=T)

# 2015-2016
demographics15 <- read.xport(here("data","DEMO_I_2015_2016.XPT"))
phthalates15 <- read.xport(here("data","PHTHTE_I_2015_2016.XPT"))
nhanes15 <- merge(phthalates15,demographics15,all.x=T)
```

## Add a column with the year to each data file:

```{r}
nhanes01$year <- 2001
nhanes03$year <- 2003
nhanes05$year <- 2005
nhanes07$year <- 2007
nhanes09$year <- 2009
nhanes11$year <- 2011
nhanes13$year <- 2013
nhanes15$year <- 2015
```

## Choose the variables to select from each data file

Key step to ensure proper weighting =
- persWeight = person-level weight ("WTINT2YR")
- psu = primary sampling unit ("SDMVPSU")
- strata = strata-level sampling unit ("SDMVSTRA")

```{r}
var<-c('year','WTINT2YR','SDMVPSU','SDMVSTRA','DMDEDUC3','DMDEDUC2','DMDHREDU','DMDHSEDU','RIDAGEYR','RIAGENDR','RIDRETH1','INDFMPIR','DMDYRSUS','DMDCITZN','URXMEP')
nhanes01s<-subset(nhanes01,select=var)
nhanes03s<-subset(nhanes03,select=var)
nhanes05s<-subset(nhanes05,select=var)
nhanes07s<-subset(nhanes07,select=var)
nhanes09s<-subset(nhanes09,select=var)
nhanes11s<-subset(nhanes11,select=var)
nhanes13s<-subset(nhanes13,select=var)
nhanes15s<-subset(nhanes15,select=var)
```

## Create one dataset for NHANES demographics and phthalates, year 2001-2016
```{r}
fullNHANES <- rbind(nhanes01s,nhanes03s,nhanes05s,nhanes07s,nhanes09s,nhanes11s,nhanes13s,nhanes15s)
```

## Save a data file. This one has what the "december22" one has, but also includes the person-level weight, primary sampling unit, and strata-level sampling unit
```{r}
## create a data file so that I don't need to load, clean, and select the data again
# has all demographics and URXMEP phthalate for NHANES year 2001-2016

# write.csv(fullNHANES, "fullNHANESfeb07.csv")
```

## Copy and rename variables so they are more intuitive. 

fpl = INDFMPIR (ratio of family income to poverty,ranges from 0 to 5 (i.e., 0% - 500%))
age = RIDAGEYR (age in years at screening)
gender = RIAGENDR
persWeight = WTINT2YR (person-level weight)
psu = SDMVPSU (primary sampling unit)
strata = SDMVSTRA (strata-level sampling unit)
refED = DMDHREDU (the household reference person's education level)
refEDspouse = DMDHSEDU (the household reference person's spouse's education level)
childED = DMDEDUC3 (the highest grade level of education completed by participants 6-19 y.o.)
adultED = DMDEDUC2 (the highest grade/level of education completed by participants 20 years and older)
ethnicity = RIDRETH1 (race/Hispanic origin)
citizenship = DMDCITZN (citizenship status)
yearsUS = DMDYRSUS (length of time in U.S.)
monoEthyl = URXMEP (mono-ethyl phthalate, ng/mL)
year = year the observation took place (*this was added by me when I merged the years of data)

```{r}
fullNHANES$fpl <- fullNHANES$INDFMPIR
fullNHANES$age <- fullNHANES$RIDAGEYR
fullNHANES$gender <- fullNHANES$RIAGENDR
fullNHANES$persWeight <- fullNHANES$WTINT2YR
fullNHANES$psu <- fullNHANES$SDMVPSU
fullNHANES$strata <- fullNHANES$SDMVSTRA
fullNHANES$refED <- fullNHANES$DMDHREDU
fullNHANES$refEDspouse <- fullNHANES$DMDHSEDU
fullNHANES$childED <- fullNHANES$DMDEDUC3
fullNHANES$adultED <- fullNHANES$DMDEDUC2
fullNHANES$ethnicity <- fullNHANES$RIDRETH1
fullNHANES$citizenship <- fullNHANES$DMDCITZN
fullNHANES$yearsUS <- fullNHANES$DMDYRSUS
fullNHANES$monoEthyl <- fullNHANES$URXMEP
```

## Create another data file that includes all the renaming and selected 30 variables

```{r}
write.csv(fullNHANES, "fullNHANES_renamed.csv")
```

## Re-categorize the variables:

#### Household reference person's education level
```{r}
fullNHANES$refED = factor(ifelse(fullNHANES$DMDHREDU %in% 1:4,"partial college and below",
                         ifelse(fullNHANES$DMDHREDU==5,"college and beyond",NA)),
                  levels=c("partial college and below","college and beyond"))

summary(fullNHANES$refED)
# 784 NAs

```

#### Household reference person's spouse's education level
```{r}
fullNHANES$refEDspouse = factor(ifelse(fullNHANES$DMDHSEDU %in% 1:4,"partial college and below",
                         ifelse(fullNHANES$DMDHSEDU==5,"college and beyond",NA)),
                  levels=c("partial college and below","college and beyond"))
summary(fullNHANES$refEDspouse)
# 10,778 NAs

```

## Highest grade level of education completed by participants 6-19 y.o.
```{r}
fullNHANES$childED = factor(ifelse(fullNHANES$DMDEDUC3 %in% 0:8,"primary",
                         ifelse(fullNHANES$DMDEDUC3 %in% 9:15,"secondary",NA)),
                  levels=c("primary","secondary"))
summary(fullNHANES$childED)
# 14,755 NAs

```


## Highest grade/level of education completed by participants 20 years and older
```{r}
fullNHANES$adultED = factor(ifelse(fullNHANES$DMDEDUC2 ==1, "less than 9th grade",
                                       ifelse(fullNHANES$DMDEDUC2 ==2, "9-11th grade",
                                       ifelse(fullNHANES$DMDEDUC2 ==3, "high school grad/GED", 
                                       ifelse(fullNHANES$DMDEDUC2 ==4, "some college or AA",
                                       ifelse(fullNHANES$DMDEDUC2 ==5, "college grad or above", NA))))))
summary(fullNHANES$adultED)
# 8,273 NAs

```


## Age in years at screening
Based on the summary statistics of the variable "RIDAGEYR": the youngest participant is 3 years old, the oldest participant is 85 years old, and the median age is 31 years old. 

```{r}
fullNHANES$age = factor(ifelse(fullNHANES$RIDAGEYR >=65,"older adult",
                    ifelse(fullNHANES$RIDAGEYR >25 & fullNHANES$RIDAGEYR<=64,"middle-aged",
                    ifelse(fullNHANES$RIDAGEYR >=19 & fullNHANES$RIDAGEYR<=25,"young adult",
                    ifelse(fullNHANES$RIDAGEYR <=18,"child",NA)))),
             levels=c("child","young adult", "middle-aged", "older adult"))
summary(fullNHANES$age)
# 220 NAs

```

## Gender
```{r}
fullNHANES$gender = factor(ifelse(fullNHANES$RIAGENDR ==1, "male",
                       ifelse(fullNHANES$RIAGENDR ==2, "female", NA)),
                levels=c("male", "female"))
summary(fullNHANES$gender)

```

## Race/ethnicity
```{r}
fullNHANES$ethnicity = factor(ifelse(fullNHANES$RIDRETH1 ==1, "Mexican American",
                       ifelse(fullNHANES$RIDRETH1 ==2, "Other Hispanic",
                       ifelse(fullNHANES$RIDRETH1 ==3, "Non-Hispanic White", 
                       ifelse(fullNHANES$RIDRETH1 ==4, "Non-Hispanic Black",
                       ifelse(fullNHANES$RIDRETH1 ==5, "Other or Multi", NA))))),
                levels=c("Non-Hispanic White", "Non-Hispanic Black", "Mexican American", "Other Hispanic", "Other or Multi"))
summary(fullNHANES$ethnicity)

```

## Citizenship Status
```{r}
fullNHANES$citizenship = factor(ifelse(fullNHANES$DMDCITZN ==1, "birth or naturalization",
                       ifelse(fullNHANES$DMDCITZN ==2, "not U,S, citizen", NA)),
                levels=c("birth or naturalization", "not U,S, citizen"))
summary(fullNHANES$citizenship)
# 42 NAs

```

## Length of time in the U.S.
```{r}
fullNHANES$yearsUS = factor(ifelse(fullNHANES$DMDYRSUS ==1, "less than 1 year",
                            ifelse(fullNHANES$DMDYRSUS ==2, "1 year or more, but less than 5 years",
                            ifelse(fullNHANES$DMDYRSUS ==3, "5 years or more, but less than 10 years",
                            ifelse(fullNHANES$DMDYRSUS ==4, "10 years or more, but less than 15 years",
                            ifelse(fullNHANES$DMDYRSUS ==5, "15 years or more, but less than 20 years",
                            ifelse(fullNHANES$DMDYRSUS ==6, "20 years or more, but less than 30 years",
                            ifelse(fullNHANES$DMDYRSUS ==7, "30 years or more, but less than 40 years",
                            ifelse(fullNHANES$DMDYRSUS ==8, "40 years or more, but less than 50 years",
                            ifelse(fullNHANES$DMDYRSUS ==9, "50 years or more", NA))))))))))
summary(fullNHANES$yearsUS)
# 18,035 NAs
 
```

## Ratio of Family Income to Poverty Level
```{r}
fullNHANES$fpl = factor(ifelse(fullNHANES$INDFMPIR < 1,"at poverty threshold",
                                        ifelse(fullNHANES$INDFMPIR >=1 & fullNHANES$INDFMPIR <2, "family income 2x poverty threshold",
                                        ifelse(fullNHANES$INDFMPIR >=2 & fullNHANES$INDFMPIR <3, "family income 3x poverty threshold",
                                        ifelse(fullNHANES$INDFMPIR >=3 & fullNHANES$INDFMPIR <4, "family income 4x poverty threshold",
                                        ifelse(fullNHANES$INDFMPIR >=4 & fullNHANES$INDFMPIR <5, "family income 5x poverty threshold",
                                        ifelse(fullNHANES$INDFMPIR ==5, "family income more than 5x poverty threshold", NA)))))),
                                 levels=c("at poverty threshold", "family income 2x poverty threshold", "family income 3x poverty threshold", "family income 4x poverty threshold", "family income 5x poverty threshold", "family income more than 5x poverty threshold"))

summary(fullNHANES$fpl)
# 1,750 NAs
```

## Save a dataset with renamed, recategorized variables

```{r}
write.csv(fullNHANES, "fullNHANES_recat.csv")
```


## Visualize Missing Data
```{r}
vis_miss(fullNHANES, sort_miss = TRUE)
```
- there are 22,349 observations
- 30 variables
- we do NOT to use these variables because lots of data is MISSING
    - "yearsUS" >> 80% 
    - "childED" >> 66%
    - "refEDspouse" >> 48%
    - "adultED" >> 37%
- all the other variables have 8% or less missing
- would the "missing" number change if I looked at (for example) childED with only the subset of participants that are in fact children?


